---
title: 'Class 15: Intro to Bioinformatics'
author: "Caitlin Johnson"
date: "2/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q1: What are those 4 candidate SNPs?
[HINT, you will may want to check the first few links of search result] 
A: rs12936231, rs8067378, rs9303277, and rs7216389 
Q2: What three genes do these variants overlap or effect?
[HINT, you can find the information from the ENSEMBLE page as shown in the image
below with red rectangles]
ZPBP2, GSDMB, and ORMDL3

## Analyze RNA-seq by genotype results 

Here we read the rssults of RNA-Seq for about 230 samples (people) with different genotypes for the rs8067378 Asthma associated SNP.

```{r}
expr <- read.table('ExpressedGenotypeResults.txt')
head(expr)
```

Extract the A/A rows 
```{r}
inds <- expr$geno == 'A/A'
```
Get the mean, median, etc of the expression of the A/A rows
```{r}
expr[inds, ]$exp
summary(expr[inds, ]$exp)
```

Get the mean, median, etc of the expression of the G/G rows
```{r}
inds2 <- expr$geno == 'G/G'
summary(expr[inds2, ]$exp)
```
Boxplot of expression of our gene of interest in individuals with different genotypes
```{r}
boxplot(exp ~ geno, data = expr, notch = T)
# if the notches don't overlap, that's a sign of possible statistical significance, but you should still do statistical tests. 
```

