---
title: 'Class 12: Structural Bioinformatics Part 1'
author: "Caitlin Johnson"
date: "2/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##PDB statistics
Here we are going to inspect the types of structures in the main database for 3D biomolecular data - the PDB.
Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. Also can
you determine what proportion of structures are protein? 
```{r}
#read in file
stats <- read.csv('Data Export Summary.csv', row.names = 1)
stats
#calculate percentage values for each 'method'
ans <- stats$Total/sum(stats$Total) *100
names(ans) <- rownames(stats)
round(ans, 2)
```
```{r}
round(sum(stats$Proteins)/sum(stats$Total) * 100, 2)
```
Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?
A: 1289
Q3: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule
in this structure?
A: Because we have it set to only show molecules of a certain size of bigger, and hydrogen is too small to cross this threshold
Q4: There is a conserved water molecule in the binding site. Can you identify this water
molecule? What residue number does this water molecule have (see note below)?
A: I think it's 308?
```{r}
library(bio3d)
lbio3d()
example("plot.bio3d")
```

###3.3 Working with Individual PDB files
```{r}
pdb <- read.pdb('1hsg')
pdb
```

Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues?
A: There are 198 amino acid residues, and the two nonprotein residues are HOH (127) and MK1 (1)
```{r}
attributes (pdb)
```


```{r}
head(pdb$atom)
#Print a subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x", "y", "z")]
```
```{r}
#Note that individual $atom records can also be accessed like this 
pdb$atom$elety[1:2]
#Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab = "B-factor")
```
Q7. What type of R object is pdb$atom? 
```{r}
str(pdb$atom)
```

A: A dataframe
```{r}
#Print a summary of the coordinate data in $xyz
pdb$xyz
```
```{r}
#Examine the row and column dimensions
dim(pdb$xyz)
```
```{r}
#Print coordinates for the first two atoms
pdb$xyz [1, atom2xyz(1:2)]
```


