---
title: 'Class 13/12B: Structural Bioinformatics II'
author: "Caitlin Johnson"
date: "2/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-Pr receptor structure
Here we will download the 1HSG PDB structure from the onine database, remove the bound drug from Merk and write out a protein only and ligand only new set of PDB files. 

We will use the functions `get.pdb()`, `read.pdb()`, `atom.select()`, and finally `write.pdb()`.
```{r}
library(bio3d)
file.name <- get.pdb('1hsg')
hiv <- read.pdb(file.name)
hiv
```
Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 
A: HOH 127, and MK1 1. Resid corresponds to water and the merk residues present in the model.
```{r}
```

### 1.2 Prepare initial protein and ligand input files
Select Protein Only
```{r}
prot <- atom.select(hiv, "protein", value = T)

write.pdb(prot, file = '1hsg_protein.pdb')
```
Select and write out ligand only file
```{r}
lig <- atom.select (hiv, "ligand", value = T )
write.pdb(lig, file = '1hsg_ligand.pdb')
```
```{r}
lig
```


