---
title: 'Class 6: R Functions'
author: "Caitlin Johnson"
date: "1/23/2020"
output: 
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions

## Level 2 Heading

### Level 3 Heading

Lets see more about **file import** (i.e. reading files into R). The main read function in base R is `read table()`


```{r}
read.csv('test1.txt')
```
For test1, we can use **read.csv** because the set-up of the table matches the **read.csv** format. We can also use **read.table**, like always, but we have to specify header and separation.

```{r}
read.table('test2.txt', header = T, sep = '$')
```
```{r}
read.table('test3.txt', sep = '')

```

#Back to Functions

Our first example function: 
```{r}
add <- function(x, y=1) {
 # Sum the input x and y
 x + y
}
```

Let's try using this function
```{r}
add(10, -2)
```
How does this work with vectors?
```{r}
add(c(1, 4, 7))
```
```{r}
add(c(1, 3, 8), 3)
```
```{r}
add(c(1, 3, 7), c(3, 2, 6))
```
```{r}
#add(c(1, 5, 8), c(2, 4))
```
This gives an error because the vectors are of different lengths
```{r}
#add(c(1, 3, 5), 2, 3, 4)
```
This also gives an error,this time because 3, 4 are unused arguments


What is the `range()` function?
```{r}
x <- c(4, 4, 10, 3, 11)
max(x)
min(x)
```
```{r}
range(x)
```

This is our secnod function
```{r}
rescale <- function(x) {
 rng <-range(x)
 (x - rng[1]) / (rng[2] - rng[1])
}
```
```{r}
rescale(x)
```

Do some more testing
```{r}
rescale(1:10)
```
```{r}
rescale(c(1,2, NA, 3, 10))
```

```{r}
x <- c(1, 2, NA, 3, 10)
rng <- range(x, na.rm = T)
rng
```

```{r}
rescale2 <- function(x) {
 rng <-range(x, na.rm = T)
 (x - rng[1]) / (rng[2] - rng[1])
}
```

```{r}
rescale2(c(1, 2, NA, 3, 10))
```

Going too far!
```{r}
rescale3  <- function(x, na.rm=TRUE, plot=FALSE) {
 rng <-range(x, na.rm=na.rm)
 print("Hello")
 answer <- (x - rng[1]) / (rng[2] - rng[1])
 print("is it me you are looking for?")
 if(plot) {
 plot(answer, typ="b", lwd=4)
 }
 print("Don't sing again!")
 return(answer)
}

```
```{r}
rescale3(x)
```
```{r}
rescale3(x, plot = T)
```

```{r}
library(bio3d)
```
```{r}
# Can you improve this analysis code?
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s3, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

```{r}
s1
```


```{r}
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.chainA
```

>Q1. What type of object is returned from the read.pdb() function? UNFINISHED

```{r}
read.pdb()
```

```{r}
s1$seqres
```
It takes one a
> What does trim.pdb() function do?

```{r}
?trim.pdb

```
It trims it down into a smaller object.

>Q4: 
A graph that contained all 3 plots on top of each other in different colors so it is easier to visualize differences between the individual plots

```{r}
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
points(s2.b, typ= 'l', col = 'blue', lwd = 2)
points(s3.b, typ = 'l', col = 'red', lwd = 2)
```

```{r}
hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

s1.b and s3.b are more similar to each other in their B-factor trends. UNFINISHED

>Q6

```{r}

improved <- function(x) {
      x.chainA <- trim.pdb(x, chain="A", elety="CA")
      x.b <- x.chainA$atom$b
      plotb3(x.b, sse = x.chainA, typ="l", ylab="Bfactor")
}
```
```{r}
improved(s1)
```

(Below) To turn in
```{r}
improved <- function(x) {
  temp <- read.pdb(x)
      temp.chainA <- trim.pdb(temp, chain="A", elety="CA")
      temp.b <- temp.chainA$atom$b
      plotb3(temp.b, sse = temp.chainA, typ="l", ylab="Bfactor")
}
```
```{r}
improved('4AKE')
```

